The polynomial representation of the type $A_{n - 1}$ rational Cherednik algebra in characteristic $p \mid n$

### 2015-05-15 formalwriteup.tex
Rewrite beginning of introduction

In this paper we study lowest-weight representations of rational Cherednik algebras associated to the symmetric group $S_n$ in characteristic $p$ dividing $n$. This problem is an interesting test case for moving results from characteristic $0$ to positive characteristic. Cherednik algebras, originally double affine Hecke algebras, were introduced by Cherednik in 1993; for an introduction to these algebras we refer to \cite{EM}. The representation theory of these algebras has been studied more extensively in characteristic $0$ than in positive characteristic, including the case of Cherednik algebras of type $A_{n-1}$ (associated to the symmetric group $S_n$). In general new techniques are required to study positive characteristic as opposed to characteristic 0. In this paper we directly connect the characteristic 0 case for the polynomial representation to the positive characteristic case; the irreducible representation of the Cherednik algebra in this case is the quotient of the polynomial representation by an ideal generated by complex residues. In the positive characteristic case, we instead let the generators be the coefficients of formal power series. In particular our argument that this quotient is a complete intersection is directly analogous to the corresponding argument in \cite{CE}, though with additional `error terms' that emerge because we must truncate the power series so that we can define it in positive characteristic. We are able to truncate the power series because we realize the $(n-1)$-dimensional representation $\h$ of $S_n$ as a quotient by the sum of the basis elements rather than as the reflection representation.

While we only consider the case of the symmetric group $S_n$ in this paper, this technique could also be used for the general case of complex reflection groups. %this bit needs revisions

%mention previous results [BEG],[CE],[GGOR]? ask
%deformation of DAHAs introduced by Etingof-Ginzburg in 2001

>>>>

The present work presents a detailed study of the polynomial representation of the type $A_{n - 1}$ rational Cherednik algebra over a field of characteristic $p$ dividing $n$.  Rational Cherednik algebras were introduced by Etingof-Ginzburg in \cite{EG} as a rational degeneration of the double affine Hecke algebra dependent on two parameters $\hbar$ and $c$.  In characteristic $0$, their type $A$ representation theory has been the subject of extensive study.  We refer the reader to \cite{EM} for a survey of these results.

In characteristic $p$ and especially in the modular case, much less is known about the representation theory of the rational Cherednik algebra.  In this paper, we consider the modular case $p \mid n$.  For $\hbar = 1$ and generic $c$, we provide a complete characterization of the irreducible quotient of the polynomial representation.  We give explicit generators for the unique maximal ideal, show that the irreducible quotient is a complete intersection, and compute its Hilbert series.

Our techniques are inspired by taking characteristic $p$ analogues of results about Cherednik algebras in characteristic $0$.  In particular, our explicit expression for generators of the maximal ideal was obtained by converting expressions containing complex residues to equivalent expressions dealing only with formal power series which may be interpreted in characteristic $p$.  While we restrict our study to type $A$ in this paper, we view this problem as a test case this philosophy and believe it may admit wider application.

In the remainder of this introduction, we give precise statements of our results and explain how they relate to other recent work.
